---
title: "viz_and_eda"
subtitle: "Part 1"
output: github_document
---

```{r setup}
library(tidyverse)
library(ggridges)
```

## Load the weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Scatterplots

Create my first scatterplot 

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + geom_point()

# new method same plot 
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()

# new method same plot
plot_weather = 
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) 

plot_weather + geom_point()
```

### Advanced Scatterplot 
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name))

# add a smooth curve and make points a bit transparent 
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE)

# lets facet some things
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  # facet_grid(name ~ .)
  facet_grid(. ~ name)

ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .2) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

## Learning Assessment 
Write a code chain that starts with weather_df; focuses only on Central Park, converts temperatures to Fahrenheit, makes a scatterplot of min vs. max temperature, and overlays a linear regression line (using options in geom_smooth()).
```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  mutate(
    tmax_fahr = tmax * (9 / 5) + 32,
    tmin_fahr = tmin * (9 / 5) + 32) %>% 
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point(alpha = .5) + 
  geom_smooth(method = "lm", se = FALSE)
```

## Odds and Ends 
```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE) 

ggplot(weather_df, aes(x = tmax, y = tmin)) + 
  geom_hex()
```

geom_hex()
geom_bin2d()
geom_density2d() 

## Learning Assessment 
In the preceding, we set the alpha aesthetic “by hand” instead of mapping it to a variable. This is possible for other aesthetics too. To check your understanding of this point, try to explain why the two lines below don’t produce the same result:

```{r}
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin), color = "blue")
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin, color = "blue"))
```

## Univariate Plots 

```{r}
ggplot(weather_df, aes(x = tmax)) + 
  geom_histogram()

ggplot(weather_df, aes(x = tmax)) + 
  geom_histogram() +
  facet_grid(. ~ name)
  
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)

# The position = "dodge" places the bars for each group side-by-side, but this gets sort of hard to understand. I often prefer density plots in place of histograms.

ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")
```

The adjust parameter in density plots is similar to the binwidth parameter in histograms, and it helps to try a few values. 

I set the transparency level to .4 to make sure all densities appear. You should also note the distinction between fill and color aesthetics here. 

You could facet by name as above but would have to ask if that makes comparisons easier or harder. 

Lastly, adding geom_rug() to a density plot can be a helpful way to show the raw data in addition to the density.

## Boxplots
```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + geom_boxplot()

ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), alpha = .5) + 
  stat_summary(fun = "median", color = "blue")

ggplot(weather_df, aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85)
```

## Learning Assessment 
Make plots that compare precipitation across locations. Try a histogram, a density plot, a boxplot, a violin plot, and a ridgeplot; use aesthetic mappings to make your figure readable
```{r}
ggplot(weather_df, aes(x = prcp)) + 
  geom_density(aes(fill = name), alpha = .5) 

ggplot(weather_df, aes(x = prcp, y = name)) + 
  geom_density_ridges(scale = .85)

ggplot(weather_df, aes(y = prcp, x = name)) + 
  geom_boxplot() 

weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, y = name)) + 
  geom_density_ridges(scale = .85)

```

## Saving and Embedding Plots 
```{r}
weather_plot = ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) 

ggsave("weather_plot.pdf", weather_plot, width = 8, height = 5)
```

